Imputation of Incomplete Multilevel Data with \pkg{mice}

### 2021-11-05 Manuscript/Manuscript.Rmd

`impact` is traumatic brain injury data with patients clustered in studies, $n_{\text{participants}} = 11022$ and $n_{\text{clusters}} = 15$, on the following 11 variables:
    * `name` Name of the study,
    * `type` Type of study (RCT: randomized controlled trial, OBS: observational cohort),
    * `age` Age of the patient,
    * `motor_score` Glasgow Coma Scale motor score,
    * `pupil` Pupillary reactivity,
    * `ct` Marshall Computerized Tomography classification,
    * `hypox` Hypoxia (0=no, 1=yes),
    * `hypots` Hypotension (0=no, 1=yes),
    * `tsah` Traumatic subarachnoid hemorrhage (0=no, 1=yes),
    * `edh` Epidural hematoma (0=no, 1=yes),
    * `mort` 6-month mortality (0=alive, 1=dead).

```{r impact}
# load data
data("impact")
# # descriptive statistics
# by(impact, impact$name, summary) 
# psych::describe(impact)[,c(2:5,8:9)]
# missingness
md_pat(impact)
```
-> Why are there no missings? According to the [vignette](https://cran.r-project.org/web/packages/metamisc/metamisc.pdf), the data is already imputed (Steyerberg et al, 2008).

@@

`popNCR` is a simulated dataset with pupils clustered in classes, $n_{\text{participants}} = 2000$, $n_{\text{clusters}} = 100$, on 7 variables: 
    * `pupil`	Pupil number within class,
    * `class`	Class number,
    * `extrav`	Pupil extraversion,
    * `sex`	Pupil gender,
    * `texp`	Teacher experience (years),
    * `popular`	Pupil popularity,
    * `popteach`	Teacher popularity.

```{r pop}
# load data
pop <- readRDS("../Data/popNCR.RDS")
# missingness
md_pat(pop)
```

@@

- Step 0: study as a predictor, AKA multilevel imputation for dummies. Doesn't work for syst missing.

## Conditional models

- How to define the imputation model(s) in `mice`?

- What do the different implementations look like?

- Step 1: Intercept

- Step 2: Slope

@@

>>>>

- What does the data look like? `impact` is traumatic brain injury data with patients clustered in studies, $n_{\text{participants}} = 11022$ and $n_{\text{clusters}} = 15$, on the following 11 variables:

  - `name` Name of the study,
  - `type` Type of study (RCT: randomized controlled trial, OBS: observational cohort),
  - `age` Age of the patient,
  - `motor_score` Glasgow Coma Scale motor score,
  - `pupil` Pupillary reactivity,
  - `ct` Marshall Computerized Tomography classification,
  - `hypox` Hypoxia (0=no, 1=yes),
  - `hypots` Hypotension (0=no, 1=yes),
  - `tsah` Traumatic subarachnoid hemorrhage (0=no, 1=yes),
  - `edh` Epidural hematoma (0=no, 1=yes),
  - `mort` 6-month mortality (0=alive, 1=dead).

```{r impact, echo=FALSE}
# load data
data("impact")
# # descriptive statistics
# by(impact, impact$name, summary) 
# psych::describe(impact)[,c(2:5,8:9)]
# missingness
plot_md_pat(impact)
```
-> Why are there no missings? According to the [vignette](https://cran.r-project.org/web/packages/metamisc/metamisc.pdf), the data is already imputed (Steyerberg et al, 2008).

@@

- What does the data look like? `popNCR` is a simulated dataset with pupils clustered in classes, $n_{\text{participants}} = 2000$, $n_{\text{clusters}} = 100$, on 7 variables:

  - `pupil`	Pupil number within class,
  - `class`	Class number,
  - `extrav`	Pupil extraversion,
  - `sex`	Pupil gender,
  - `texp`	Teacher experience (years),
  - `popular`	Pupil popularity,
  - `popteach`	Teacher popularity.

```{r pop, echo=FALSE, message=FALSE, warning=FALSE}
# load data
pop <- readRDS("../Data/popNCR.RDS")
# ICCs
icc_pop <- multilevel::ICC1(aov(popular ~ as.factor(class), data = pop))
icc_teach <- multilevel::ICC1(aov(popteach ~ as.factor(class), data = pop))
```
- What are the ICCs? For `popular` the ICC is `r round(icc_pop, 2)`. For `popteach` it is `r round(icc_teach, 2)`.

- What does the missingness look like? Induced MAR/MNAR missingness. Missing data pattern:

```{r pop_pat, echo=FALSE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
# missingness
plot_md_pat(pop)
```

- Does the missing data of popular depend on popteach? One could for example check this by making a histogram of popteach separately for the pupils with known popularity and missing popularity. We do see that the histogram for the missing popular (TRUE) is further to the right than the histogram for observed popular (FALSE). This would indicate a right-tailed MAR missingness. In fact this is exactly what happens, because we created the missingness in these data ourselves. But we can make it observable by examining the relations between the missingness in popular and the observed data in popteach.

- Does the missingness in teacher popularity depend on pupil popularity? Yes: there is a dependency. The relation seems to be right-tailed.

```{r pop_dist, echo=FALSE, message=FALSE, warning=FALSE}
# distributions
plot_NA_cond(pop, x = "popteach", z = "popular", bins = 1)
plot_NA_cond(pop, x = "popular", z = "popteach", bins = 1)
```

@@

- Step 0: As predictor

- Step 1: Random intercepts

- Step 2: Random slopes

@@

- What do the different implementations look like? How to define the imputation model(s) in `mice`?

## Step 0

- AKA multilevel imputation for dummies.

- Doesn't work for systematic missingness.

## Step 1-3

- TODO: fill in.
